\chapter{Fixed Income Securities}
A fixed income security is an investment that provides a return in the form of fixed periodic
interest payments and the eventual return of principal at maturity. The payments of a fixed income
security are known in advance. Bonds are the most common form of fixed income securities.

A derivative is an instrument which value is derived from a primary instrument (stocks, currency,
bonds, other derivatives etc.) also called underlying instrument. Derivative contracts divided into
OTC-derivatives (over the counter), stock exchange derivatives. Derivatives are priced using
Martingale pricing theory. Derivatives are effective tools for controlling risk. With them you can
separate risk and trade it separately. Enabeling traders to sell unwanted risk or buy risk to
diversify the portfolio.

\section{Forward contract}
A forward contract is an agreement between a buyer and a seller to trade an underlying asset at a
future date for a price specified at the contract's inception. In order to have a single
transaction, the price of a forward is such that the value is zero for both parties (Most common:
Currency forwards, forward rate agreements.)

\subsection{Pricing}
A forward with maturity $T$ and delivery price $K$ provides a payoff of $P_{T}-K$ at time $T$ and no
other payments. Here, $P$ is the underlying (typically the price of an asset or a specific interest
rate). In an arbitrage free market, there exists an equivalent martingale measure $\Q$ (associated with the money market numeraire), such that the time $t<T$ value of this payoff is:
\[
V_{t}=\E_{t}^{\Q}\left[\e^{-\int_{t}^{T} r_{u} \d 
u}\left(P_{T}-K\right)\right]=\E_{t}^{\Q}\left[\e^{-\int_{t}^{T} r_{u} \d 
u} P_{T}\right]-K B_{t}^{T}
,\]
where $B_t^{T}$ denotes the time $t$ price of a zero-coupon bond which pays 1 at maturity $T$. 

For $t \le T$, the forward price is set to the value of $K$ that solves $V_{t}=0$, since this
implies that neither party must make an upfront payment. Thus,
\[
F_{t}^{T}=\frac{\E_{t}^{\Q}\left[\e^{-\int_{t}^{T} r_{u} \d  u} P_{T}\right]}{B_{t}^{T}}
\]
Under the risk-neutral measure $\Q$, the discounted asset price is a martingale.
So for an asset with no intermediate income before $T$ we have $F_{t}^{T}=\frac{P_{t}}{B_{t}^{T}}$.

\section{European options}
European options provide the right, but not the obligation to trade the underlying asset on
expiration date for a pre-specified (at time $t$) price $K$ (American options can be exercised on
any date). A Call is the right to buy, and a Put is the right to sell the asset.

\subsection{Pricing}
Consider a call option with expiration date $T$ and strike $K$ on the underlying $P_{t}$. The payoff
at time $T$ (and hence the call price at $T$) is
\[
C_{T}=\max \left\{P_{T}-K, 0\right\}
\]
Again, in an arbitrage free market, there exists an equivalent martingale measure $\Q$, such that
the discounted asset prices are martingales. 

Assuming that the underlying and zero-coupon bonds are traded, we may change the $\Q$-measure to the
$T$-forward martingale measure $\Q^{T}$, which uses $B_t^{T}$ as numeraire.
Using this measure we can derive a general pricing formula for a European call option:
\begin{align*}
C_{t} & =B_{t}^{T} \E_{t}^{\Q^{T}}\left[\max \left\{P_{T}-K, 0\right\}\right] \\
& =B_{t}^{T} \E_{t}^{\Q^{T}}\left[\1_{\left\{P_{T}>K\right\}}\left(P_{T}-K\right)\right] \\
& =B_{t}^{T}\left(\E_{t}^{\Q^{T}}\left[\1_{\left\{P_{T}>K\right\}} P_{T}\right]-K \Q^{T}_t\left(P_{T}>K\right)\right) \\
& =B_{t}^{T} \E_{t}^{\Q^{T}}\left[\1_{\left\{P_{T}>K\right\}} P_{T}\right]-K B_{t}^{T} \Q^{T}_t\left(P_{T}>K\right) \\
\text{(Change of numeraire)}\quad & =\E_{t}^{\Q^{P}}\left[B_{t}^{T}P_T \frac{B_{T}^{T}}{P_T}
\frac{P_t}{B_{t}^{T}} \1_{\left\{P_{T}>K\right\}}\right]-K B_{t}^{T} \Q_{t}^{T}\left(P_{T}>K\right) \\
\text{(Measurable)}\quad & =P_{t} \E_{t}^{\Q^{P}}\left[\1_{\left\{P_{T}>K\right\}}\right]-K B_{t}^{T} \Q_{t}^{T}\left(P_{T}>K\right) \\
& =P_{t} \Q_{t}^{P}(P_{T}>K)-K B_{t}^{T} \Q_{t}^{T}(P_{T}>K)
\end{align*}
where $\Q^{P}$ is the martingale measure using $P_{t}$ as numeraire, thus we have assumed that $P_t$
is also traded in the market. The proof is completely analogous for European puts, only we use the
payoff $\max \{ K - P_T, 0 \}$. Note that the final probabilities correspond to the option being in
the money under the two different equivalent martingale measures. To compute these probabilities
further assumptions of the market model is needed, for instance if we set
\[ 
\Q^{P}_t (P_T > K) = N(d_1) \quad  \text{ and } \quad 
\Q^{T}_t (P_T > K) = N(d_2)
,\]
where $N(\cdot )$ denotes the standard normal cumulative distribution function. The general
pricing formula reduces to the Black-Scholes call option price.


\section{Relevant results}


\subsection{Bonds, Yields and Rates}

\fxnote{Fill out this section}
$\quad $

\subsection{Other Fixed-Income securities}


\begin{description}
\item[Interest rate forwards, forward rate agreements (FRAs)] 
\end{description}


\fxnote{Add pricing formulas}
\begin{description}
\item[Future] 
\end{description}
Like a Forward, a Futures contract is a contract where the two parties agree to trade the underlying
at a perspeciefed strike. The main difference is that futures are marked-to-market, at least daily.
This means that the two counterparties continuously settle the difference between the previous
futures price and the new future price. 

Thus, you are implicitly exposed to the current interest rate, through the future price. 
Whereas, with a forward you only take into account the expected future rate. 
Thus, the difference in price between the two contracts is the discounted covariance between the
zero-coupon bond and the payment price at time $T$. A Futures contract has value: 
\[
  \phi_{t}^{T}=\E^{\Q}\left[P_{T}\right]
.\]

\begin{description}
\item[Futures on bonds] 
\end{description}

\begin{description}
\item[Options on bonds] 
\end{description}


\begin{description}
\item[Caps] 
\end{description}
A cap protects a floating-rate borrower against the risk of paying very high interest rates.
Work as a ceiling on the interest rate, compensating the holder if the rate rises above a specified
level. Consists of a series of underlying options called caplets (call option on the spot interest
rate).

\begin{description}
\item[Floors] 
\end{description}
A floor protects a floating-rate lender against the risk of receiving very low interest
rates. Works like caps, consists of floorlets (put option on floating interest rate with delayed
payment of the payoff).

\begin{description}
\item[Collars] 
\end{description}
A combination of caps and floors. Ensures that the interest rate payments on a floating
rate borrowing arrangement stay between two prespecified levels. (Can be seen as a portfolio of a
long position in a cap with rate $K_{c}$ and a short position in floor with rate $K_{f}$.)

\begin{description}
\item[Swaps] 
\end{description}
An interest rate swap is an exchange of two cash flow streams determined by certain interest
rates. A plain vanilla interest rate swap takes two forms:

\begin{itemize}
  \item Payer swap: Pays a stream of fixed rate payments and recieves a stream of floating rate
    payments.
  \item Receiver swap: Pays a stream of floating rate payments and receives a stream of fixed rate
    payments.
\end{itemize}
Forward swap: Agreement to enter into a swap with a future starting date and fixed rate already
set.

\begin{description}
\item[Forward swaps] 
\end{description}
An agreement to enter a swap at an agreed-upon future starting date $T_0$ and fixed rate. The
forward swap rate $\tilde{L}_t^{\delta, T_0}$ is the fixed rate that solves $P_t = 0$ at time $t\le
T$.

\begin{description}
\item[Swaptions] 
\end{description}
A European swaption gives its holder the right at the expiry date to enter into an
interest rate swap that has a given fixed rate. A payer swaption gives the right to enter into a
payer swap, and likewise for receiver swaption.

\begin{description}
\item[Repurchase agreements] 
\end{description}

Short term contracts used for lending money based on high-grade securities as collateral (government
bonds). Low interest rate, but also (a bit) risky, as is might be difficult to renew loans in a
distressed economy (think fall 2008, illuquidity).

\begin{itemize}
\item Parties agree on Prices $P_0$ and $P_T = P_0 (1+rT)$
\item At time $t=0$, the security is traded at price $P_0$, and the security repurchased at time
  $T$, using price $P_T$
\end{itemize}
The collateral reduces the lender’s exposure to counter party (credit) ris.  Repos are subject to
daily margin calls if collateral value changes (like futures) 

Financial crisis: some bank could not ''repay'' the repo due to a cut in funding by another bank
(Goldman Sachs, GS), so not enough capital to do so. GS bought credit default swaps a few days
before, so probably knew what was going to happen.










\fxnote{Maybe a section on the different martingale measures we have seen}
